Place of death in NRS COVID-19 deaths: data processing for analysis

### 2021-10-19 covid_nrs_place_of_death_data.Rmd
exported proportions by place and geography over time file

```{r}
proportions_of_deaths_by_place_by_geography <-
  bind_rows(
    ## historic deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      past_deaths_hb %>% mutate(geography = "hb"),
      past_deaths_la %>% mutate(geography = "la")
    ) %>%
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
    ,
    ## current deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      total_deaths_hb %>% mutate(geography = "hb"),
      total_deaths_la %>% mutate(geography = "la")
    ) %>%
    select(year, place_of_death, number_of_deaths = deaths_all_causes, geography, ref_area) %>%  # keep only relevant columns & rename deaths due to all causes
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
  )
```

>>>>

```{r}
proportions_of_deaths_by_place_by_geography <-
  bind_rows(
    ## historic deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      past_deaths_hb %>% mutate(geography = "hb"),
      past_deaths_la %>% mutate(geography = "la")
    ) %>%
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
    ,
    ## current deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      total_deaths_hb %>% mutate(geography = "hb"),
      total_deaths_la %>% mutate(geography = "la")
    ) %>%
    select(year, place_of_death, number_of_deaths = deaths_all_causes, geography, ref_area) %>%  # keep only relevant columns & rename deaths due to all causes
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
  )

## write to csv
proportions_of_deaths_by_place_by_geography %>%
  write_csv(file = paste0(dir_outputs,"/proportions_of_deaths_by_place_by_geography.csv"))
```
